---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# DiffScan

<!-- badges: start -->
<!-- badges: end -->

DiffScan is a computational framework for differential analysis of RNA structure probing experiments at nucleotide resolution. Specifically, it identifies RNA structurally variable regions (SVRs) between two cellular conditions.

DiffScan is compatible with various RNA structure probing platforms.

## System Requirements

`DiffScan` is built and tested upon [R](https://cran.r-project.org/) (version 4.0.2). It mainly depends on the following R packages.

``
dplyr,
rlist,
pipeR,
pbmcapply,
MASS,
coin,
plyr,
ggplot2,
seqinr,
SpatialExtremes,
preprocessCore
``



## Installation

Install `DiffScan` in [R](https://cran.r-project.org/) or [Rstudio](https://www.rstudio.com/products/rstudio/) with:

``` r
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
if (!requireNamespace("preprocessCore", quietly = TRUE))
    install.packages("http://bioconductor.org/packages/3.11/bioc/src/contrib/preprocessCore_1.50.0.tar.gz")
    
devtools::install_github("yub18/DiffScan")
```

Typical install time: a few minutes or less.

## Usage

Suppose there are two RNA transcripts (`t1, t2`) with structure probing data in condition A and B, with 2 replicates in each condition (`A1, A2, B1, B2`). To identify SVRs in `t1, t2`, first format the data like the following.

```{r example}
library(DiffScan)
library(dplyr)
set.seed(123)
r = list(
  t1 = tibble(A1 = runif(50), A2 = A1 + runif(50, -0.1, 0.1),
              B1 = A1 + runif(50, -0.1, 0.1), B2 = B1 + runif(50, -0.1, 0.1)),
  t2 = tibble(A1 = runif(150), A2 = A1 + runif(150, -0.1, 0.1),
              B1 = c(A1[1:90] + runif(90, -0.1, 0.1), rep(0, 10), 
                     A1[101:130] + runif(30, -0.1, 0.1), rep(1, 20)), 
              B2 = B1 + runif(150, -0.1, 0.1))
)
```

In this example, transcript `t1` has no SVRs, and transcript `t2` has 2 SVRs covering nucleotide positions 91 nt ~ 100 nt and 131 nt ~ 150 nt.

Second, initialize `DiffScan` as follows. (See the document of `DiffScan::init` for appropriate quality control.)

```{r}
r_init = init(r, qc_ctrl = F, ncores = 1)
```

Third, perform normalization to remove systematic bias in reactivities. (Users convinced of the comparability between reactivities can skip this step.)

```{r, results='hide'}
r_norm = normalize(r_init, ncores = 1)
```

Next, run the Scan algorithm to identify SVRs.

```{r}
res = scan(r_norm$r, seed = 123, alpha = 0.05, ncores = 1)
```

`DiffScan` predicts no SVRs for transcript `t1`, and predicts 2 SVRs for transcript `t2` covering nucleotide positions 132 nt ~ 150 nt and 92 nt ~100 nt.

```{r}
res
```

Expected run time: a few minutes or less.

## Citation
